Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-25 draft.tex
Minor modifs

\begin{document}
\maketitle
\begin{abstract}
Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters).
In the classical setting, the voting rule is fixed beforehand and full preference information is provided.
Recently, the assumption of full preference information has been questioned by a number of researchers and several methods for eliciting preferences have been proposed.

@@

\section{Introduction}
%\commentOC{Do we agree that we should talk systematically about agents (or systematically about voters)? Currently we talk about voters in first section, then agents.}
Aggregation of preference information is a central task in many computer systems (recommender systems, search engines, etc).
In many situations, such as in group recommender systems, the goal is to find a consensus choice.
It is therefore natural to look at methods from social choice and see how they can be adapted for group decision-making in a computerized setting.

@@

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights\footnote{We assume all positions from $1$ to $m-1$ may contribute positively to the score of an alternative. The case where it is known that only the first $k<m-1$ positions matter can be easily dealt with by specifying $w_r = 0$ for $r > k$ as constraints. }:
% If the chair considers that, for example, “only the first ten positions matter” (so that $w_r = 0$ for $r > 10$), then we simply assume the chair is able to express this constraint at the start of the querying process, thus, that this information is known, and approximate the problem by considering that it is only about the first ten alternatives, thus, set $m=10$.}
\begin{align}
1=w_{1} > w_{2} > \ldots > w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$. 
%\commentOC{I suggest we say no more than this to justify this assumption. Otherwise, we might say something like the following.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. 
The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference in scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of score is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
In what follows let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
%There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies 
\end{claim}
%In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.
\noindent It follows that $\hat{v}_{j}(x)$, the rank of $x$ for voter $j$ in the PMR-maximizing linear order $\bar{v}_{j}$ %$\PMR(x,y; \pprofile, W)$ 
is $\hat{v}_{j}(y) = 1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $\hat{v}_{j}(x)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\beta = 0$ otherwise.

@@

%In summary, some best completion $\pref$ satisfies two properties: (1) for all $c \pinc x$, $c \pref x$; and (2) for all $c \pinc y$, if not $c \pref x \pref y$ then $y \pref c$. 
%These two properties are intuitively understandable by observing that maximizing the PMR requires placing $x$ as low as possible, hence the first property; and $y$ as high as possible, hence the second property, with the exception that sometimes it is not possible to satisfy both goals and noting that the first goal is more important by convexity of the weights. 
%A proof that this reasoning is correct is in \cref{sec:prfCompl}.
\begin{claim}
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W)  
%\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \\ 
& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
\end{align}
where the $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{claim}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
%We now show that the pairwise maximum regret can be computed with a linear program.
%The PMR can be now computed by optimizing a linear program (LP). 
%Constraints are on the feasible parameters (the weights of the scoring rules).
The last proposition shows us that PMR is linear in the weights.
We remind that the preferences of the chair are encoded with linear constraints $\Co_{W}$.
% contraints represent the preferences $\Co_{W}$ of the chair about the scoring rule.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can be obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$ (weights attached to the different positions). 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& \Co_W
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).

@@

This concludes the proof of the following claim. \commentOC{I don’t think the reorganization with moving the claim slightly upwards works. The proof really starts much higher and would be cut in half if moving the claim slightly. Moving it more would require more reorganisation of the text.}
\begin{claim}
	Given a rational $\lambda$ and a rank $r$ between $1$ and $m - 2$, the profile named $P'$ in the foregoing description is such that, for any weight vector $\w \in \W$, $a \in f(P')$ iff $w_{r} - w_{r+1} ≥ \lambda (w_{r+1} - w_{r+2})$ and $b \in f(P')$ iff $w_{r} - w_{r+1} ≤ \lambda (w_{r+1} - w_{r+2})$, where $f$ is the PSR parameterized with $\w$.
\end{claim}

@@

\paragraph{Elicitation strategies}
\commentOC{TODO move this.} At this stage, we are content that a procedure exists to transform our questions to questions about winners of a PSR. Further studies would investigate, for example, what is lost in terms of elicitation efficiency when we are forced to restrain to realistic concrete questions, meaning, questions involving small profiles; or investigate the relationship between our understanding of scores as (cardinal) utility and (ordinal) scores as definition of a PSR.

We develop several elicitation strategies for simultaneous elicitation of agents' preferences and of the scoring rule.
While it is of course possible to first fully elicit the agents’ preferences and afterwards elicit weights, we also want to propose interleaved approaches and compare them experimentally.
Indeed, it can be beneficial to interleave questions asked to the chair and questions asked to agents, depending on which is estimated to be more informative.

@@

\commentOC{TODO Continue here.}
Two phase method.
Ask a predefined (non adaptive) sequence of questions in order to learn the weights $w$ of the scoring rule).
Then use minimax regret as in Lu and Boutilier.

%\item Non-interleaved regret method: comparison with a method based on minimax regret, but not interleaved. 
\medskip \noindent
{\em Non-interleaved non-regret strategies (useful as benchmarks)}	
	\begin{itemize}
		\item Random strategy: randomly chooses an agent $i$ and a comparison query such that $x \pinc_i y$.
		\item {\em Volumetric} strategy: chooses an agent $i$ and a query that maximizes the number of new pairwise preferences revealed given the worst response.
	\end{itemize}

% unless $\MMR=0$ we ask a question that reduce uncertainty

>>>>

\begin{document}
\maketitle
\begin{abstract}
Social choice deals with the problem of determining a consensus choice from the preferences of different agents (voters).
In the classical setting, the voting rule is fixed beforehand and full information concerning the preferences of the agents is provided.
Recently, the assumption of full preference information has been questioned by a number of researchers and several methods for eliciting preferences have been proposed.

@@

\section{Introduction}
Aggregation of preference information is a central task in many computer systems (recommender systems, search engines, etc).
In many situations, such as in group recommender systems, the goal is to find a consensus choice.
It is therefore natural to look at methods from social choice and see how they can be adapted for group decision-making in a computerized setting.

@@

\medskip
We also assume that the weights of the scoring rule are only partially specified.
Therefore the vector $(w_1,\ldots,w_m)$  is not known but we are given a set of constraints restraining the possible values that weights can take.
We consider a decreasing sequence of weights\footnote{We assume all positions from $1$ to $m-1$ may contribute positively to the score of an alternative. The case where it is known that only the first $k<m-1$ positions matter can be easily dealt with by specifying $w_r = 0$ for $r > k$ as constraints. } \commentOC{It’s not clear to me that this is true…}:
% If the chair considers that, for example, “only the first ten positions matter” (so that $w_r = 0$ for $r > 10$), then we simply assume the chair is able to express this constraint at the start of the querying process, thus, that this information is known, and approximate the problem by considering that it is only about the first ten alternatives, thus, set $m=10$.}
\begin{align}
1=w_{1} > w_{2} > \ldots > w_{m}=0. \label{eq:monotone}
\end{align}
This is a natural assumption, as it is better to be ranked first than second, second than third, etc. 
Without loss of generality, we consider that $w_1=1$ and $w_m=0$. 
%\commentOC{I suggest we say no more than this to justify this assumption. Otherwise, we might say something like the following.

@@

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. 
The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
%There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies 
\end{claim}
%In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.
It follows that $\hat{v}_{j}(x)$, the rank of $x$ in the PMR-maximizing linear order of voter $j$, %$\PMR(x,y; \pprofile, W)$ 
is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\beta = 0$ otherwise.

@@

%In summary, some best completion $\pref$ satisfies two properties: (1) for all $c \pinc x$, $c \pref x$; and (2) for all $c \pinc y$, if not $c \pref x \pref y$ then $y \pref c$. 
%These two properties are intuitively understandable by observing that maximizing the PMR requires placing $x$ as low as possible, hence the first property; and $y$ as high as possible, hence the second property, with the exception that sometimes it is not possible to satisfy both goals and noting that the first goal is more important by convexity of the weights. 
%A proof that this reasoning is correct is in \cref{sec:prfCompl}.

\begin{claim}
The $\PMR$ can be written as:
\begin{align} 
\PMR(x,y; \pprofile, W)  
%\max_{w \in W} s(y; \hat{\profile}, \w) - s(y; \hat{\profile}, \w)\\
& = \max_{w \in W} \sum_{j=1}^n w_{\hat{v}_j(y)} - w_{\hat{v}_j(x)} = \\ 
& = \max_{w \in W} \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_i. 
\end{align}
where the $\hat{\alpha}_{r}^{y}$ (resp. $\hat{\alpha}_{r}^{x}$)  is the number of times $y$ (resp. $x$) is at rank $r$ in the rankings of the completed profile $\hat{\profile}$.
%$\hat{\alpha}_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
\end{claim}
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
%We now show that the pairwise maximum regret can be computed with a linear program.
%The PMR can be now computed by optimizing a linear program (LP). 
%Constraints are on the feasible parameters (the weights of the scoring rules).
The last proposition shows us that PMR is linear in the weights.
We remind that the preferences of the chair are encoded with linear constraints $\Co_{W}$.
% contraints represent the preferences $\Co_{W}$ of the chair about the scoring rule.
The pairwise max regret $\PMR(x,y; \pprofile,W)$ can be obtained as the solution of the following linear program defined on the variables $w_1, …, w_m$ (weights attached to the different positions). 
\begin{align*}
\max_{\w} & \sum_{r=1}^m (\hat{\alpha}_{r}^{y} - \hat{\alpha}_{r}^{x}) w_{r}\\
\text{ s.t. } & \text{Equation } (\ref{eq:monotone})\\
& \text{Equation } (\ref{eq:convexity})\\
& \Co_W
\end{align*}
Note that given our choice $w_{1}=1$ and $w_{m}=0$, there are only $m-2$ variables 
(we leave $w_{1}$ and $w_{m}$ in the LP just for clarity of presentation).

@@

This concludes the proof of the following claim. 
\begin{claim}
	Given a rational $\lambda$ and a rank $r$ between $1$ and $m - 2$, the profile named $P'$ in the foregoing description is such that, for any weight vector $\w \in \W$, $a \in f(P')$ iff $w_{r} - w_{r+1} ≥ \lambda (w_{r+1} - w_{r+2})$ and $b \in f(P')$ iff $w_{r} - w_{r+1} ≤ \lambda (w_{r+1} - w_{r+2})$, where $f$ is the PSR parameterized with $\w$.
\end{claim}
\commentOC{Plz others check this.} At this stage, we are content that a procedure exists to transform our abstract questions to questions about winners of a PSR. Further studies would investigate, for example, what is lost in terms of elicitation efficiency when we are forced to restrain to realistic concrete questions, meaning, questions involving small profiles; or investigate the relationship between our understanding of scores as (cardinal) utility and (ordinal) scores as definition of a PSR.

@@

\paragraph{Elicitation strategies}
We develop several elicitation strategies for simultaneous elicitation of agents' preferences and of the scoring rule.
While it is of course possible to first fully elicit the agents’ preferences and afterwards elicit weights, we also want to propose interleaved approaches and compare them experimentally.
Indeed, it can be beneficial to interleave questions asked to the chair and questions asked to agents, depending on which is estimated to be more informative.

@@

\commentOC{TODO continue from here.}
The \strat{Two phase} strategy asks a predefined, non adaptive sequence of questions in order to gather informations about the weights $w$ of the scoring rule. \commentOC{To check.}
Then use minimax regret as in Lu and Boutilier.

%\item {\em Volumetric} strategy: chooses an agent $i$ and a query that maximizes the number of new pairwise preferences revealed given the worst response.
